### 2015-11-14 use-cases/niche-modelling.Rmd

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

In this example, we plot actual occurrence data for _Bradypus_ species against a single predictor variable, BIO1 (annual mean temperature). This is only ont step in a species distribution modelling nworkflow.

This example can be done using BISON data as well with our rbison package.

_Load libraries_

```{r}
library("rgbif")
library("dismo")
library("maptools")
library("plyr")
```

_Raster files_

Make a list of files that are installed with the dismo package, then create a rasterStack from these

```{r}
files <- list.files(paste(system.file(package = "dismo"), "/ex", sep = ""),
                    "grd", full.names = TRUE)
predictors <- stack(files)
```

_Get world boundaries_

```{r}
data(wrld_simpl)
```

_Get GBIF data using the rOpenSci package rgbif_

```{r cache=TRUE}
nn <- name_lookup("bradypus*", rank = "species")
nn <- na.omit(unique(nn$data$nubKey))
df <- occ_search(taxonKey = nn, hasCoordinate = TRUE, limit = 500)
df <- df[ sapply(df, function(x) class(x$data)) %in% "data.frame" ]
df <- ldply(lapply(df, "[[", "data"))
df2 <- df[,c('decimalLongitude','decimalLatitude')]
```

_Plot_

(1) Add raster data, (2) Add political boundaries, (3) Add the points (occurrences)

```{r fig.keep='last'}
plot(predictors, 1)
plot(wrld_simpl, add = TRUE)
points(df2, col = "blue")
```