### 2015-11-14 use-cases/oak-occurrences.Rmd

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

This example comes from [Antonio J. Perez-Luque](https://twitter.com/ajpelu) who [shared his plot on Twitter][tweet]. Antonio compared the occurrences of Valley Oak (_Quercus lobata_) from [GBIF][gbif] to the distribution of the same species from the [Atlas of US Trees][atlas].

_Load libraries_

```{r}
library('rgbif')
library('raster')
library('sp')
library('maptools')
library('rgeos')
library('scales')
```

_Get GBIF Data for Quercus lobata_

```{r cache=TRUE}
keyQl <- name_backbone(name='Quercus lobata', kingdom='plants')$speciesKey
dat.Ql <- occ_search(taxonKey=keyQl, return='data', limit=50000)
```

_Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)_

From [http://esp.cr.usgs.gov/data/little/](http://esp.cr.usgs.gov/data/little/). And save shapefile in same directory

```{r}
url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
tmp <- tempdir()
download.file(url, destfile = "~/querloba.zip")
unzip("~/querloba.zip", exdir = tmp)
ql <- readShapePoly(file.path(tmp, "querloba.shp"))
```

_Get Elevation data of US_

```{r}
alt.USA <- getData('alt', country = 'USA')
```

_Create Hillshade of US_

```{r}
alt.USA <- alt.USA[[1]]
slope.USA <- terrain(alt.USA, opt = 'slope')
aspect.USA <- terrain(alt.USA, opt = 'aspect')
hill.USA <- hillShade(slope.USA, aspect.USA, angle = 45, direction = 315)
```

_Plot map_

```{r fig.keep='last'}
plot(hill.USA, col = grey(0:100/100), legend = FALSE, xlim = c(-125, -116), ylim = c(32, 42), main = 'Distribution of Quercus lobata', xlab = "Long", ylab = 'Lat')
# add shape from Atlas of US Trees
plot(ql, add = TRUE, col = alpha("white", 0.6), border = FALSE)
# add Gbif presence points
points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = -121, y = 40.5, "GBIF Data", pch = 19, col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
legend(x = -121, y = 41.5, "Atlas of United States Trees \n (Little, E. 1971)", pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
```

[tweet]: https://twitter.com/ajpelu/status/473951167567757312
[gbif]: http://www.gbif.org/
[atlas]: http://esp.cr.usgs.gov/data/little/