### 2016-03-11 manuscript.Rmd
converted ms to discuss python in addition to R client

---
layout: review, 11pt
linenumbers: true
title: "R and Python clients for GBIF species occurrence data"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: scott(at)ropensci.org
    footnote: Corresponding author
  - name: Carl Boettiger
    affiliation: boettig
    email: carl(at)ropensci.org
address:
  - code: cstar
    address: |
      University of California, Berkeley, CA, USA
  - code: boettig
    address: |
      University of California, Berkeley, CA, USA
abstract: |
      1. xxx

      2. xxx

      3. xxx

      4. xxxx

bibliography: components/references.bib
csl: components/peerj.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(
  tidy = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = 1,
  comment = '#>',
  collapse = TRUE,
  verbose = TRUE
)
basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
```

# Introduction

Perhaps the most fundamental element in many fields of ecology is
the individual organism. The number of individuals of each species in
a given location forms the basis for many sub-fields of ecology and
evolution. Some research questions necessitate collecting new data, while
others can easily take advantage of existing data. In fact, some
ecology fields are built largely on existing data, e.g., macro-ecology
[@Brown_1995; @Beck_2012].

Data on individuals, including which species, and where they're found,
can be used for a large number of research questions. Biodiversity
records have been used for a suite of other use cases:
validating habitat suitability models with real occurrence data
[@Ficetola_2014]; ancestral range reconstruction [@Ferretti_2015;
@Mendoza_2015]; development of invasive species watch lists
[@Faulkner_2014]; evaluating risk of invasive species spread
[@Di_Febbraro_2013]; and effects of climate change on future
biodiversity [@Brown_2015].

In addition to wide utility, this data is important for conservation.
Biodiversity loss is one of the greatest challenges of our time
[@Pimm_2014], and some have called this the sixth great mass
extinction [@Ceballos_2015]. Given this challenge there is a great
need for data on specimen records, whether collected from live sightings
in the field or specimens in museums.

# Global Biodiversity Information Facility

There are many online services that collect and maintain specimen records.
However, Global Biodiversity Information Facility (hereafter, GBIF,
[http://www.gbif.org](http://www.gbif.org)) is the largest collection of
biodiversity records globally, currently with 643 million records, 1.6 million
taxa, 15,450 datasets from 780 publishers (as of 2016-02-09).
Many large biodiversity warehouses such as iNaturalist
([http://www.inaturalist.org](http://www.inaturalist.org)), VertNet
([http://vertnet.org](http://vertnet.org)), and USGS's Biodiversity Information
Serving Our Nation (BISON; [http://bison.usgs.ornl.gov](http://bison.usgs.ornl.gov))
all feed into GBIF.

The most important data organizational level in GBIF is the individual record. In R
using `rgbif` you'll recognize this as a row in the data.frame that is returned from
`occ_search()` or `occ_data()`. Going upstream, each record is part of a dataset,
where each dataset is submitted by an organization, organizations are organized into
nodes, datasets are published through institutions (which may be hosted at another
organization), and a network is a group of datasets (managed by GBIF).

Each record has some taxonomic name associated with it, which itself is linked to
a lot of other taxonomic data. GBIF maintains their own taxonomic data.

All this organization matters because you can navigate through GBIF data through all
these organizational levels.

# The clients

## The rgbif package

Herein, we describe the `rgbif` software package [@rgbif] for working with
GBIF data in the R programming environment [@R]. R is a widely used
language in academia, as well as non-profit and private sectors.
Importantly, R makes it easy to execute all steps of the
research process, including data management, data manipulation
and cleaning, statistics, and visualization. Thus, an R client
for getting GBIF data is a powerful tool to facilitate
reproducible research.

```{r echo=FALSE, results='hide', eval=FALSE}
install.packages("rgbif", dependencies=TRUE)
```

```{r echo=FALSE, results='hide'}
library("rgbif")
```

The `rgbif` package is nearly completely written in R (a small Javascript
library is included for reading well known text [@wkt]), uses an [MIT license][mit]
to maximize use everywhere. `rgbif` is developed publicly on GitHub
at [https://github.com/ropensci/rgbif][ghrgbif], where development versions
of the package can be installed, and bugs and feature requests reported.
Stable versions of `rgbif` can be installed from [CRAN][cranrgbif], the
distribution network for R packages. `rgbif` is part of the rOpenSci project
(http://ropensci.org), a developer network making R software to facilitate
reproducible research.

## The pygbif library

`pygbif` [@pygbif] is a Python library for working with GBIF data in
the Python programming environment [@python]. Python is a general
purpose programming language used widely in all sectors, and for
all parts of software development including server and client
side use cases. Python is used exclusively in some scientific
disciplines (e.g., astronomy), and has partial usage in other
disciplines. A Python client for GBIF data is an important tool
given the even wider usage of Python than R, though maybe slightly
less than R for ecology/biology.

```{python eval=FALSE}
pip install pygbif
```

```{python eval=FALSE}
import pygbif
```

The `pygbif` library is less mature and complete than the R package. It also
uses an [MIT license][mit] to maximize use everywhere. `pygbif` is developed
publicly on GitHub at [https://github.com/sckott/pygbif][ghpygbif], where
development versions of the package can be installed, and bugs and feature
requests reported. Stable versions of `pygbif` can be installed from
[pypi][pypipygbif], the distribution network for Python libraries.

### Library interfaces

`rgbif` and `pygbif` are designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, taxonomic names,
occurrences, and maps. We also include functions to interface with the OAI-PMH
GBIF service; only dataset information is available vis this service, however. We
ignore maps in both libraries as it is concerned with generating maps primarily for web
applications. Both libraries have a suite of functions dealing with each of registry, taxonomic
names, and occurrences - we'll go through each in turn describing design and example usage.

## GBIF headers

With each request `rgbif` and `pygbif` make to GBIF's API, we send request headers that
tell GBIF what client the request is coming from, including what version of
the library. This helps GBIF know what proportion of requests are coming from
which client, and therefore from R vs. Python; this information is helpful for GBIF
in thinking about how people are using GBIF data.

## Registry

The GBIF registry API services are spread across five sets of functions
via the main GBIF API:

* Datasets
* Installations
* Networks
* Nodes
* Organizations

Dataset information in general is available via the OAI-PMH service,
functions in `rgbif` prefixed with `gbif_oai_`, but not available in `pygbif`
yet.

Datasets are owned by organizations. Organizations are endorsed by nodes to share
datasets with GBIF. Datasets are published through institutions, which may be
hosted at another organization. A network is a group of datasets (managed by GBIF).
Datasets are the units that matter the most with respect to registry information,
while installations, networks, nodes, and organizations are simply higher level
organizational structure.

### Datasets

Dataset functions include search, dataset metadata retrieval, and dataset metrics.
Searching for datasets is an important part of the discovery process. One can
search for datasets on the GBIF web portal. However, programmatic searching using
this package is much more powerful. Identifying datasets appropriate for a
research question is helpful as you can get metadata for each dataset, and track
down dataset specific problems, if any.

The `dataset_search()` function is one way to search for datasets. Here, we search
for the query term "oregon", which finds any datasets that have terms matching
that term.

```{r}
res <- dataset_search(query = "oregon")
res$data$datasetTitle[1:10]
```

See also `datasets()` and `dataset_suggest()` in `rgbif` for searching for datasets.

In Python, we can similary search for datasets. Here, search for datasets of type
`OCCURRENCE`:

```{python eval=FALSE}
from pygbif import registry
registry.datasets(type="OCCURRENCE")
```

#### Dataset metrics

Dataset metrics are another useful way of figuring out what
datasets you may want to use. One drawback is that these metrics data are only
available for datasets of type _checklist_, but there are quite a lot of
them (`r dataset_search(type = "checklist")$meta$count`).

Here, we search for dataset metrics for a single dataset, with uuid
`ec93a739-1681-4b04-b62f-3a687127a17f`, a checklist of the ants (Hymenoptera:
Formicidae) of the World.

```{r eval=FALSE}
res <- dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
data.frame(rank = names(res$countByRank),
           count = unname(unlist(res$countByRank)))
```

```{r echo=FALSE}
res <- dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
df <- data.frame(rank = names(res$countByRank),
                 count = unname(unlist(res$countByRank)))
knitr::kable(df)
```

And in Python, get metrics for the same dataset as above:

```{python}
from pygbif import registry
registry.dataset_metrics(uuid='ec93a739-1681-4b04-b62f-3a687127a17f')
```

### Networks, nodes, and installations

Networks, nodes and installations are at a higher level of organization
above datasets, but can be useful if you want to explore data from given
organizations. Here, in R we search for the first 10 GBIF networks, returning
just the title field.

```{r}
networks(limit = 10)$data$title
```

And in Python:

```{python}
from pygbif import registry
registry.networks(limit = 10)
```

## Taxonomic names

The GBIF taxonomic names API services are spread across five functions in `rgbif`:

* Search GBIF name backbone - `name_backbone()`
* Search across all checklists - `name_lookup()`
* Quick name lookup - `name_suggest()`
* Name usage of a name according to a checklist - `name_usage()`
* GBIF name parser - `parsenames()`

`pygbif` only has `name_backbone()` and `name_suggest()` at this time.

The goal of these name functions is often to settle on a taxonomic name
known to GBIF's database. This serves two purposes: 1) when referring to
a taxonomic name, you can point to a URI on the Internet, and 2) you can
search for metadata on a taxon, and occurrences of that taxon in
GBIF.

Taxonomic names are particularly tricky. Many different organizations have
their own unique codes for the same taxonomic names, and some taxonomic
groups have preferred sources for the definitive names for that group. That's
why it's best to determine what name GBIF uses, and its associated identifier,
for the taxon of interest instead of simply searching for occurrences with a
taxonomic name.

When searching for occurrences (see below) you can search by taxonomic
name (and other filters, e.g., taxonomic rank), but you're probably better off
figuring out the taxonomic key in the GBIF backbone taxonomy, and using
that to search for occurrences. The `taxonkey` parameter in the GBIF occurrences
API expects a GBIF backbone taxon key.

### GBIF Backbone

The GBIF backbone taxonomy is used in GBIF to have a consistent way to refer
to taxonomic names throughout their services. The backbone has `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$distinctNamesCount` unique
names and `r dataset_metrics("d7dddbf4-2cf0-4f39-9b2a-bb099caae36c")$countByRank$SPECIES` species names. The backbone taxonomy is also a dataset with key
`d7dddbf4-2cf0-4f39-9b2a-bb099caae36c`
(http://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c).

We can search the backbone taxonomy with the function `name_backbone()` in both
R and Python clients. Here, we're searching for the name _Poa_, restricting to
genera, and the family _Poaceae_, in R

```{r}
res <- name_backbone(name='Poa', rank='genus', family='Poaceae')
res[c('usageKey', 'kingdom')]
```

and in Python

```{python eval=FALSE}
from pygbif import species
res = species.name_backbone(name='Poa', rank='genus', family='Poaceae')
[ res[x] for x in ['usageKey', 'kingdom'] ]
```

### Name searching

One of the quickest ways to search for names is using `name_suggest()`, which
does a very quick search and returns minimal data. Here, we're searching for
the query term _Pum_, and we get back many names:

```{r eval=FALSE}
name_suggest(q='Pum', limit = 6)
```

```{r echo=FALSE}
df <- name_suggest(q='Pum', limit = 6)
knitr::kable(df)
```

The same in Python

```{python eval=FALSE}
from pygbif import species
species.name_suggest(q='Pum', limit = 6)
```

With these results, you can then proceed to search for occurrences with the
taxon key(s), or drill down further with other name searching functions to
get the exact taxon of interest.

## Occurrences

GBIF provides two ways to get occurrence data: through the
`/occurrence/search` route (see `occ_search` in `rgbif`, or
`occurrences.search` in `pygbif`), or via the
`/occurrence/download` route (many functions, see below).
`occ_search()`/`occurrences.search` is the main function for the search route,
and is more appropriate for smaller data, while
the download functions are more appropriate for larger
data requests.

GBIF imposes for any given search a limit of 200,000 records in the search
service, after which point you can't download any more records for that
search. However, you can download more records for different searches.

We think the search service is still quite useful for many people
even given the 200,000 limit. For those that need more data, we have
created a similar interface in the download functions, that should
be easy to use. Users should take note that using the download service
has a few extra steps to get data into R, but is straight-forward.

The download service, like the occurrence search service, is rate-limited.
That is, you can only have one to three downloads running simultaneously for
your user credentials. However, simply check when a download job is complete,
then you should be able to start a new download request.

### Download API

The download API syntax is similar to the occurrence search API
in that the same parameters are used, but the way in which the query
is defined is different. For example, in the download API you can
do greater than searches (i.e., `latitude > 50`), whereas you cannot
do that in the occurrence search API. Thus, unfortunately, we couldn't
make the query interface exactly the same for both search and
download functions.

Using the download service can consist of as few as three steps: 1) Request
data via a search; 2) Download data; 3) Import data into R.

Request data download given a query. Here, we search for the taxon key
`3119195`, which is the key for _Helianthus annuus_
(http://www.gbif.org/species/3119195).

```{r eval = FALSE}
occ_download('taxonKey = 3119195')
#> <<gbif download>>
#>   Username: xxxx
#>   E-mail: xxxx
#>   Download key: 0000840-150615163101818
```

You can check on when the download is ready using the functions
`occ_download_list()` and `occ_download_meta()`. When it's ready
use `occ_download_get()` to download the dataset to your computer.

```{r}
(res <- occ_download_get("0000840-150615163101818", overwrite = TRUE))
```

What's printed out above is a very brief summary of what was downloaded,
the path to the file, and its size (in human readable form).

Next, read the data in to R using the function `occ_download_import()`.

```{r eval=FALSE}
library("dplyr")
dat <- occ_download_import(res)
dat %>%
  select(gbifID, decimalLatitude, decimalLongitude)
#>       gbifID abstract accessRights accrualMethod accrualPeriodicity accrualPolicy alternative audience
#> 1  725767384       NA                         NA                 NA            NA          NA       NA
#> 2  725767447       NA                         NA                 NA            NA          NA       NA
#> 3  725767450       NA                         NA                 NA            NA          NA       NA
#> 4  725767513       NA                         NA                 NA            NA          NA       NA
#> 5  725767546       NA                         NA                 NA            NA          NA       NA
#> 6  725767579       NA                         NA                 NA            NA          NA       NA
#> 7  725767609       NA                         NA                 NA            NA          NA       NA
#> 8  725767645       NA                         NA                 NA            NA          NA       NA
#> 9  725767678       NA                         NA                 NA            NA          NA       NA
#> 10 725767681       NA                         NA                 NA            NA          NA       NA
#> ..       ...      ...          ...           ...                ...           ...         ...      ...
#> Variables not shown: available (lgl), bibliographicCitation (chr), conformsTo (lgl), contributor (lgl),
#>      coverage (lgl), created (chr), creator (lgl), date (lgl), dateAccepted (lgl), dateCopyrighted
#>      (lgl), dateSubmitted (lgl), description (lgl), educationLevel (lgl), extent (lgl), format (lgl),
#>      hasFormat (lgl), hasPart (lgl), hasVersion (lgl), identifier (chr), instructionalMethod (lgl),
```

In Python

```{python eval = FALSE}
from pygbif import occurrences as occ
occ.download('taxonKey = 3119195')
(res = occ.download_get("0000840-150615163101818", overwrite = True))
```

We don't have `pygbif` functionality at the moment for importing data, but
it's coming soon.

#### Downloaded data format

The downloaded dataset from GBIF is a Darwin Core Archive (DwC-A), an
internationally recognized biodiversity informatics standard
([http://rs.tdwg.org/dwc/](http://rs.tdwg.org/dwc/)). The DwC-A downloaded is a
compressed folder with a number of files, including metadata, citations for each
of the datasets included in the download, and the data itself, in separate files for
each dataset as well as one single `.txt` file. In `rgbif::occ_download_import()`, we
simply fetch data from the `.txt` file. If you want to dig into the metadata,
citations, etc., it is easily accessible from the folder on your computer.

### Search API

The search API follows the GBIF API and is broken down into the following functions:

* Get a single numeric count of occurrences - `rgbif`: `occ_count()` / `pygbif`: `occurrences.count`
* Search for occurrences - `rgbif`: `occ_search()` / `pygbif`: `occurrences.count`
* A simplified and optimized version of `rgbif`: `occ_search()` or `occ_data()` / `pygbif`: `occurrences.count`
* Get occurrences by occurrence identifier - `rgbif`: `occ_get()` / `pygbif`: `occurrences.count`
* Get occurrence metadata - `rgbif`: `occ_metadata()` / `pygbif`: `occurrences.count`

#### Search for occurrences

__R__

The main search work-horse is `occ_search()`. This function allows very flexible
search definitions. In addition, this function does paging internally, making it
such that the user does not have worry about the 300 records per request limit -
but of course we can't go over the 200,000 maximum limit.

The output of `occ_search()` presents a compact `data.frame` so that no matter
how large the `data.frame`, the output is easily assessed because only a few of
the records (rows) are shown, only a few columns are shown (with others shown in
name only), and metadata is shown on top of the `data.frame` to indicate data
found and returned, media records found, unique taxonomic hierarchies returned,
and the query executed.

The output of these examples, except one, aren't shown, but all run correctly.

Search by species name, using `name_backbone()` first to get key

```{r}
(key <- name_suggest(q = 'Helianthus annuus', rank = 'species')$key[1])
occ_search(taxonKey = key, limit = 2)
```

Instead of getting a taxon key first, you can search for a name directly

```{r eval=FALSE}
occ_search(scientificName = 'Ursus americanus')
```

Search for many species

```{r eval=FALSE}
splist <- c('Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa')
keys <- sapply(splist, function(x) name_suggest(x)$key[1], USE.NAMES = FALSE)
occ_search(taxonKey = keys, limit = 5, return = 'data')
```

Spatial search, based on well known text format [@wkt], or a bounding box set of
four coordinates

```{r eval=FALSE}
# well known text
occ_search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
# bounding box
occ_search(geometry = c(-125.0,38.4,-121.8,40.9), limit = 20)
```

Get only occurrences with lat/long data

```{r eval=FALSE}
occ_search(hasCoordinate = TRUE, limit = 5)
```

Get only those occurrences with spatial issues. Spatial issues are a set of issues
that are returned in the `issues` field. They each indicate something different
about that record. For example, the issue `COUNTRY_COORDINATE_MISMATCH` indicates
that the interpreted occurrence coordinates fall outside of the indicated country.
You can see how that might be useful when it comes to cleaning your data prior
to analysis/visualization.

```{r eval=FALSE}
occ_search(hasGeospatialIssue = TRUE, limit = 5)
```

__Python__

The equivalent occurrence search workhorse in `pygibf` is `occurrences.search`.

Search by species name, using `name_backbone()` first to get key

```{python eval=FALSE}
from pygbif import species
from pygbif import occurrences as occ
key = species.name_suggest(q = 'Helianthus annuus', rank = 'species')['data'][0]['key']
occ.search(taxonKey = key, limit = 2)
```

Instead of getting a taxon key first, you can search for a name directly

```{python eval=FALSE}
occ.search(scientificName = 'Ursus americanus')
```

Search for many species

```{python eval=FALSE}
from pygbif import species
from pygbif import occurrences as occ
splist = ['Cyanocitta stelleri', 'Junco hyemalis', 'Aix sponsa']
keys = [ species.name_suggest(x)['data'][0]['key'] for x in splist ]
occ.search(taxonKey = keys, limit = 5)
```

Spatial search, based on well known text format [@wkt], or a bounding box set of
four coordinates

```{python eval=FALSE}
from pygbif import occurrences as occ
# well known text
occ.search(geometry = 'POLYGON((30.1 10.1, 10 20, 20 40, 40 40, 30.1 10.1))', limit = 20)
# bounding box
occ.search(geometry = '-125.0,38.4,-121.8,40.9', limit = 20)
```

Get only occurrences with lat/long data

```{python eval=FALSE}
from pygbif import occurrences as occ
occ.search(hasCoordinate = True, limit = 5)
```

Get only those occurrences with spatial issues. Spatial issues are a set of issues
that are returned in the `issues` field. They each indicate something different
about that record. For example, the issue `COUNTRY_COORDINATE_MISMATCH` indicates
that the interpreted occurrence coordinates fall outside of the indicated country.
You can see how that might be useful when it comes to cleaning your data prior
to analysis/visualization.

```{python eval=FALSE}
from pygbif import occurrences as occ
occ.search(hasGeospatialIssue = True, limit = 5)
```

#### Data cleaning

GBIF provides optional data issues with each occurrence record. These issues fall
into many different pre-defined classes, covering issues with taxonomic names,
geographic data, and more (see `occ_issues_lookup()` to find out more information
on GBIF issues; and the same data on [GBIF's development site][gbifissues]).

`occ_issues()` provides a way to easily filter data downloaded via `occ_search()`
based on GBIF issues.

```{r}
out <- occ_search(issue = 'DEPTH_UNLIKELY', limit = 500)
NROW(out)
out %>% occ_issues(-cudc) %>% .$data %>% NROW
```

There's no equivalent interface in `pygbif`.

## Mapping

An obvious downstream use case for species occurrence data is to map the
data. `rgbif` per se is largely not concerned with making this easier,
although we do have a simple wrapper around `ggplot2` to make it easy to
get a quick plot of occurrence data. For example, here we plot 100
occurrences for _Puma concolor_.

```{r fig.width=8, fig.height=4}
key <- name_backbone(name='Puma concolor')$speciesKey
dat <- occ_search(taxonKey = key, limit = 100, hasCoordinate = TRUE)
gbifmap(dat$data)
```

Another package, [mapr][mapr], is the perfect mapping companion to `rgbif`.
It has convenient functions for handling input data from `rgbif`, `spocc`,
or arbitrary data.frame's, and output plots for base plots, `ggplot2`,
`ggmap` (`ggplot2` with map layers underneath), and interactive maps on
GitHub gists or with Leaflet.js.

There's no equivalent interface in `pygbif`.

## GBIF data in other R packages

We discuss usage of GBIF data in other R packages throughout the
manuscript, but provide a synopsis here for clarity.

### taxize

Some of the GBIF taxonomic services are also available in [taxize][taxize],
an R package that focuses on getting data from taxonomic data sources on
the web. For example, with `get_gbifid()` one can get GBIF IDs used for
a set of taxonomic names - then use those IDs in other functions in `taxize`
to get additional information, like taxonomically downstream children.

### spocc

GBIF occurrence data is available in the R package [spocc][spocc] via `rgbif`.
`spocc` is a unified interface for fetching species occurrence data from
many sources on the web. For example, a user can collect occurrence data
from GBIF, iDigBio, and iNaturalist, and easily combine them, then use other
packages to clean and visualize the data.

# Use cases

The following are three use cases for `rgbif`: niche modeling,
spatial change in biodiversity, and distribution mapping.

### Ecological niche modeling

```{r niche-modeling, child="use-cases/niche-modelling.Rmd"}
```

### Biodiversity in big cities

```{r city-diversity, child="use-cases/city-diversity.Rmd"}
```

### Valley oak occurrence data comparison

```{r oak-occurrences, child="use-cases/oak-occurrences.Rmd"}
```

# Conclusions and future directions

The `rgbif` and `pygbif` libraries provide programmatic interfaces to GBIF's
application programming interface (API) - a powerful tool for making
research using species occurrence data reproducible. In fact,
the `rgbif` package has already been used in 22 scholarly publications
(as of 2015-11-14).

The `rgbif` package is relatively stable, and should not have
many breaking changes unless necessitated due to changes in the
GBIF API.

The `pygbif` library is still in early development, and will
greatly benefit from any feedback and use cases.

One area of focus in the future is to attempt to solve many use
cases that have been brought up with respect to GBIF data. For
example, some specimens are included in GBIF that are located
in botanical gardens. For many research questions, researchers are
interested in "wild" type occurrences, not those in human curated
scenarios. Making removal of these occurrences easy would be
very useful, but is actually quite a hard problem.

# Acknowledgments

This project was supported in part by the Alfred P Sloan Foundation
(Grant No. G-2014-13485), and in part by the Helmsley Foundation
(Grant No. 2016PG-BRI004).

# Data Accessibility

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
[github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
can be found at [https://github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
under an MIT license.

# References

[mit]: http://choosealicense.com/licenses/mit/
[ghrgbif]: https://github.com/ropensci/rgbif
[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
[gbifdocs]: http://www.gbif.org/developer/summary
[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
[taxize]: https://github.com/ropensci/taxize
[spocc]: https://github.com/ropensci/spocc
[mapr]: https://github.com/ropensci/mapr
[ghpygbif]: https://github.com/sckott/pygbif
[pypipygbif]: https://pypi.python.org/pypi/pygbif