### 2017-09-26 use-cases/oak-occurrences.Rmd
hopefully fixed issue #8

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

This example comes from [Antonio J. Perez-Luque](https://twitter.com/ajpelu) who [shared his plot on Twitter][tweet]. Antonio compared the occurrences of Valley Oak (_Quercus lobata_) from [GBIF][gbif] to the distribution of the same species from the [Atlas of US Trees][atlas].

_Load libraries_

```{r}
library('rgbif')
library('raster')
library('sp')
library('sf')
library('rgeos')
library('scales')
library('rnaturalearth')
```

_Get GBIF Data for Fraxinus excelsior_

```{r cache=TRUE}
keyFe <- name_backbone(name = 'Fraxinus excelsior', kingdom = 'plants')$speciesKey
dat.Fe <- occ_search(taxonKey = keyFe, return = 'data', limit = 10000L)
```

_Get Distribution map of F. excelsior European Forest Genetic Resources Programme_

From <http://www.euforgen.org/species/fraxinus-excelsior/>. And save shapefile in same directory

```{r cache = TRUE}
url <- 'http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus_excelsior.zip'
tmp <- tempdir()
download.file(url, destfile = "fraxinus_excelsior.zip")
unzip("fraxinus_excelsior.zip", exdir = tmp)
fe <- sf::read_sf(file.path(tmp, "Fraxinus_excelsior_EUFORGEN.shp"))
```

_Get Elevation data of US_

```{r cache = TRUE}
eur <- rnaturalearth::ne_countries(continent = "europe", type = "map_units")
eur1 <- eur[eur$sovereignt != "Russia", ]
```

_Plot map_

```{r fig.keep='last'}
plot(eur1, col = "darkgrey", legend = FALSE,
     main = 'Distribution of Fraxinus excelsior')
# add distribution range layer
plot(fe, add = TRUE, col = alpha("white", 0.5), border = FALSE)
# add Gbif presence points
points(dat.Fe$decimalLongitude, dat.Fe$decimalLatitude,
       cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = 38, y = 81, c("GBIF Data", "Range Layer"), pch = 19, bg = "grey",
       col = c('darkgreen', alpha("white", 0.5)), pt.cex = 1, cex = .8)
```

[tweet]: https://twitter.com/ajpelu/status/473951167567757312
[gbif]: http://www.gbif.org/
[atlas]: http://esp.cr.usgs.gov/data/little/